### 2015-05-14 notes.tex
initial commit

-smoothed analysis

--what is it? option when theory is far from practice, so good in the
learning setting. randomly perturb (``smooth'') part of a problem and
examine what happens on that smoothed instance. give a few classic
examples here; simplex, tsp.

--what assumptions does it make about the data? when is it useful to
apply?  when something about problem is not adversarial; i.e. may
arise randomly in practice.

--smoothed analysis in learning. still new; people unsure where to
apply it: lots of different things to smooth. smooth distribution or
smooth points? what does it mean to smooth distribution? don't
know. classically in SA there's one clear thing to smooth but in
learning we don't really know what's best yet. these 2 papers contain
different approaches. k-means is runtime-focused. smooths points and
shows that the runtime on the perturbed set is expected
polynomial. tensor decomp is accuracy focused. perturbs vectors and
shows that factorization works robustly with perturbations. these 2
approaches are just examples in the current lines of research. we
could also imagine doing [OTHER CREATIVE THING] In classic model,
algorithm performance is measured by runtime. In learning we have
another metric: accuracy. can use smoothing to get results on both
fronts; see the two examples below.

% maybe do the main result from the tensor paper instead and mention
% this result at a higher level. in the conclusion mention how the two
% papers take different avenues.

\section{$k$-means}
\subsection{Algorithm}
The $k$-means algorithm solves the \emph{clustering
  problem}, set as follows. Given a set $S$ of points in a (generally
high-dimensional) space $\mathbb{R}^d$, output a $k$-partition of the
set (a collection of $k$ ``clusters'') such that the elements in each
cluster are similar under a chosen metric. $k$-means is iterative:
first it arbitrarily chooses the $k$ cluster centers, then assigns each point in $S$ to the closest center. Next, it
loops over these clusters and computes the mean of all
points in the cluster. These means become the new cluster centers and
the assignment process repeats. The algorithm terminates when the
local improvements to cluster centers no longer affect the assignment
of points.

\subsection{$k$-means is Exponential}
In \cite{KMworstcase}, Vattani shows that an adversary can choose
initial cluster centers that will force $k$-means to iterate
$2^{\Omega(n)}$ times for $d \ge 2$. This construction is based on
[SOMETHING].

\subsection{Smoothed Analysis}
We give a summary of the argument in \cite{SAkmeans}, by Arthur,
Manthey, and Roeglin, that $k$-means has polynomial smoothed
complexity. The main result of this section is the following theorem.
\begin{theorem}
Fix a set $S^{\prime} \subset [0,1]^d$ of $n$ data points, and
independently perturb each point in $S^{\prime}$ by a Gaussian
distribution with mean 0 and variance $\sigma^2$. Label the new set
$S$. Then the expected running time of $k$-means on $S$ is bounded by
a polynomial in $n$ and $1/\sigma$.
\end{theorem}
We reason about the potential function \begin{equation}
\Psi = \sum_{x \in S}||x - c(x)||^2,
\end{equation} 
where $c(x)$ is the cluster center to which the point $x$ is assigned
in the current iteration. The proof relies on two lemmas: first, that
$\Psi$ is bounded by a polynomial after one iteration (define
iteration) of the algorithm, and second that $\Psi$ decreases by a
polynomial amount in each subsequent iteration (actually, sequences of
a few iterations).

\begin{lemma}
The potential after one iteration is bounded by a polynomial.
\end{lemma}
\begin{proof}
\end{proof}