### 2015-05-14 paper.tex

\documentclass[11pt]{article}

\usepackage{amsthm, amssymb, amsmath}
\usepackage{rotating}
\usepackage{verbatim}
\usepackage{enumerate}
\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption}
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{hyperref} % For hyperlinks in the PDF
\usepackage[shortlabels]{enumitem}
\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the ``Abstract'' text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\titlelabel{\thetitle.\quad}
%\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
%\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
%\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
%\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{May 2015} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\linespread{1.05}
\bibliographystyle{plain}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{theorem}{Theorem}

% --------------------
% TITLE SECTION
% --------------------

\title{\vspace{-5mm}\fontsize{22pt}{10pt}\selectfont Smoothed Analysis in Learning: Tensors and $k$-Means}

\author{
\large
\textsc{Hunter Lang} \\
\normalsize MIT \\
\normalsize \href{mailto:hjl@mit.edu}{\texttt{hjl@mit.edu}}\\
\and
\textsc{Carlos Cortez} \\
\normalsize MIT \\
\normalsize \href{mailto:cortezc@mit.edu}{\texttt{cortezc@mit.edu}}\\
\vspace{-5mm}
}
\date{}

\begin{document}
\maketitle

%--------------------
% ABSTRACT
%--------------------

\begin{abstract}
\noindent Most learning problems are hard in the worst-case, so much
of the current research focuses on finding good heuristics,
polynomial-time approximation algorithms, or special cases with
proveable accuracy and runtime guarantees. But many algorithms that
are inefficient in the worst case (Simplex being the classical
example) consistently seem to run fast in practice. Smoothed analysis
gives a framework for better understanding performance on real-world
data, specifically for problems where some component is not
adversarial. This is a natural assumption in learning, since data in
the learning setting are usually prone to measurement or modeling
noise. We survey the (very much ongoing) application of smoothed
analysis to learning problems by way of two examples: $k$-means and
tensor decomposition.
\end{abstract}

\section{Introduction}
Spielman and Teng \cite{SA}, \cite{SAtwo} introduced smooth analysis
to give a more suitable framework for predicting real-life algorithm
performance. Not every algorithm that runs fast in practice is
polynomial-time; worst-case analysis falls short of explaining why
some ``slow'' algorithms are actually quite efficient in practice. The
original paper \cite{SA} gave a proof that the Simplex algorithm has
polynomial ``smoothed complexity'': that is, if you assume the data
are subject to random perturbations (the kind that would arise from
measurement noise in practice), Simplex runs in expected polynomial
time. This sparked a host of smoothed analysis proofs involving
classic combinatorial optimization problems. [EXAMPLES]. The key
assumption inherent in applying smoothed analysis techniques to an
algorithm is that some component of the problem is not
adversarial. The hope is that worst-case inputs are somehow isolated
in the input space (indeed, the worst-case inputs to many algorithms
are intricate and fragile), so that any real data will, with high
probability, not be worst-case. In recent years, there has been an increasing trend of applying smoothed analysis to learning problems \cite{SAtwo}, \cite{SAkmeans}, \cite{PAC}, \cite{TD}.

\section{Tensor Decomposition}
\subsection{Algorithm}
\subsection{Worst-case}
\subsection{Smoothed Analysis}

\section{Conclusion}

\begin{thebibliography}{1}
  \bibitem{SA}
    Daniel A. Spielman and Shang{-}Hua Teng,
    ``Smoothed Analysis of Algorithms: Why the Simplex Algorithm Usually takes Polynomial Time.''
    \emph{Journal of the ACM, Vol 51 (3)},
    pp. 385 - 463,
    2004.

  \bibitem{SAtwo}
    Daniel A. Spielman and Shang{-}Hua Teng,
    ``Smoothed Analysis: An Attempt to Explain the Behavior of Algorithms in Practice''
    \emph{Communications of the ACM, Vol. 52 No. 10},
    pp. 76-84,
    2009.

  \bibitem{KMworstcase}
    Andrea Vattani,
    ``k-means Requires Exponentially Many Iterations Even in the Plane.'' 
    \emph{ Proc. of the 25th ACM Symp. on Computational Geometry (SoCG)}, 
    pp 324–332, 
    2009.

  \bibitem{SAkmeans}
    David Arthur, Bodo Manthey, and Heiko Roeglin,
    ``Smoothed Analysis of the $k$-means Method.''
    2010.

  \bibitem{PAC}

  \bibitem{TD}
    Aditya Bhaskara, Moses Charikar, Ankur Moitra, and Aravindan Vijayaraghavan,
    ``Smoothed Analysis of Tensor Decomposition'',
    CoRR,
    2013.

\end{thebibliography}
\end{document}